WritingSSfiles

### 2019-05-24 CreateStockSynthesis.Rmd
Make CreateStockSynthesis.Rmd work in more general case across species by modifing species config file

This page discusses reading in an existing stock assessment file and replacing the data values with those from the `atlantisom` package. To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`

@@

Dummy files are in the package directory; use `r4ss` to load the dummy data file

```{r readdata, echo=FALSE}
require(r4ss)
#Read in sardine data file
sardine.data <- r4ss::SS_readdat_3.30("./inst/extdata/Sardine_SS_files/sardEM_3_3.dat")
```

Below, we can see the structure of the catch and CPUE data
```{r}
names(sardine.data$catch) <- names(sardine.data$CPUE)

sardine.data$CPUE
```

We can next look at age composition data
```{r}
sardine.data$agecomp

sardine.data$lencomp
```

@@

```{r}
#CPUE units, SD and error structure
sardine.data$CPUEinfo

#Number of length bins
sardine.data$N_lbins

#Length bin method
sardine.data$Lbin_method

#Vector of age bins
sardine.data$agebin_vector

@@

```{r}
results <- get(load("outputCCV3run_truth.RData"))
devtools::load_all()
species=c("Pacific_sardine")

@@

#Efficiency of the survey
effic <- data.frame(species=species, efficiency=0.5)

# Assume uniform selectivity
sel<- data.frame(species=rep(species,10),
               agecl=c(1:10),
              selex=rep(1,10))

#Sample survey - 3rd timestep simulates a summer survey
survey_out <- create_survey(dat=results$nums, time=seq(3,495,by=5), species=species, boxes=boxes, effic=effic, selex=sel)

#Set effective sample size for age compositions
effN <- 1000
highEffN <- data.frame(species=species, effN=rep(effN, length(species)))

@@

#Extract length composition data
ss_length_stdsurv <- calc_age2length(structn = ss_structnstd,
                                     resn = ss_resnstd,
                                     nums = age_comp_data,
                                     biolprm = results$biolprm, fgs = results$fgs,
                                     CVlenage = 0.1, remove.zeroes=TRUE)

@@

# CV function
cv <- data.frame(species="Pacific_sardine", cv=0.1)

@@

```{r}
sardine.catch <-  create_fishery_subset(dat = results$catch,
                                         time = seq(1,99,by=1),
                                         species = species,
                                         boxes = boxes)

#Sample catch in numbers
sardine.catch.total <- sardine.catch %>%
  group_by(time) %>%
  summarise(catch=sum(atoutput))

nyears <- 50
burnin <- 30
fish_years <- burnin:(burnin+nyears-1)
survey_years <- seq(3,495,by=5)[burnin:(burnin+nyears-1)]

#Test writing CPUE in biomass
  sardine.data <- SS_write_ts(ss_data_list = sardine.data,
                ts_data = list(survObsBiom$atoutput[fish_years],
                  sardine.catch.total$catch[fish_years]),
                CVs = c(0.1,0.1),
                data_years = list((survObsBiom$time[fish_years]-3)/5+1,             sardine.catch.total$time[fish_years]),
            sampling_month = list(rep(7,nyears),
                                rep(1,nyears)),
                units = c("biomass","numbers"),
                data_type=c("CPUE","catch"),
                fleets = c(2,1))

@@

#Get the age bins
age_bin_names <- names(sardine.data$agecomp)[10:length(names(sardine.data$agecomp))]
age_bins <- sub("a","",age_bin_names)

@@

## Write age composition data for survey
sardine.data <- SS_write_comps(ss_data_list = sardine.data,
               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
               data_rows = list(sardine.data$styr:(sardine.data$styr+nyears-1)),
               sampling_month = list(rep(7,nyears)),
               data_type = c("agecomp"),
               fleet_number = c(1),
               bins = list(age_bins),
               caal_bool = c(FALSE))
sardine.data$agecomp

#We end up using CAAL for the survey below, so let's generate fishery age comps instead
effN <- data.frame(species="Pacific_sardine", effN=effN)

catch_numssshigh <- sample_fish(sardine.catch, effN)

@@

catch_length <- calc_age2length(structn = catch_structnss,
                                 resn = catch_resnss,
                                 nums = catch_numssshigh,
                                 biolprm = results$biolprm, fgs = results$fgs,
                                 maxbin = 200,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

@@

## Setting up sardine EM
For the sardine EM, we want conditional age-at-length (CAAL) and length composition data only. So we use the same function with our matrices of length and CAAL comp data to write to the data file.

@@

apply_month <- list(rep(7, nrow(comp_list[[1]])), rep(7, nrow(comp_list[[2]])), rep(1,nrow(comp_list[[3]])),rep(1,nrow(comp_list[[4]])))

# Write CAAL and length composition data
sardine.data <- SS_write_comps(ss_data_list = sardine.data, comp_matrix = comp_list, data_rows = list((comp_list[[1]]$time-3)/5 + 1 ,(survey_years-3)/5 + 1,fish_years,fish_years),
        sampling_month = apply_month, data_type = rep(c("agecomp", "lencomp"),2), fleet_number = c(2,2,1,1),  bins = list(age_bins, length_bins, age_bins, length_bins), caal_bool = c(TRUE, rep(FALSE,3)))

head(sardine.data$lencomp)
head(sardine.data$agecomp)

#Change length bin structure to match atlantis data
sardine.data$lbin_vector <- length_bins

#Set lbin_method to 1 - this makes the population length bins match the data bins 
#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
sardine.data$lbin_method <- 1
sardine.data$binwidth <- "#"
sardine.data$minimum_size <- "#"
sardine.data$maximum_size <- "#"

#Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
sardine.data$age_info$minsamplesize <- rep(0.001,2)

SS_writedat_3.30(sardine.data, outfile = "./inst/extdata/Sardine_SS_files/sardEM_3_3.dat",
                 overwrite=TRUE)
```

@@

@@

>>>>

This page discusses reading in an existing stock assessment file and replacing the data values with those from the `atlantisom` package. To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`. You will need to ensure the relevant Stock Synthesis files are stored in the inst/extdata/ folder, with each species example in its own folder.

@@

Dummy files are in the package directory; use `r4ss` to load the dummy data file. Use the relevant species config file to load the configuration variables for each run. Note: this script will extract conditional age-at-length data, length composition data, and a CPUE in biomass for one survey during the specified time interval, and age composition data, length composition data, and catch in numbers from a single fishery. The timing and duration of the survey and fishery should be set in the config file. Unless otherwise stated, numbers going into Stock Synthesis are in thousands of fish. Weight of a single fish is in grams or kilograms, and weight of biomass is in metric tons. We assume a single sex model with a sex ratio fixed at 0.5 - different assumptions require modifying the Stock Synthesis control file. This script only runs with Stock Synthesis 3.30.

```{r readdata, echo=FALSE}
require(r4ss)
source("./config/sardine_config.R")

stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0("./inst/extdata/",
model_dir,
datfile_name))
```

Below, we can see the structure of the catch and CPUE data
```{r}
names(stocksynthesis.data$catch) <- names(stocksynthesis.data$CPUE)

stocksynthesis.data$CPUE
```

We can next look at age composition data
```{r}
stocksynthesis.data$agecomp

stocksynthesis.data$lencomp
```

@@

```{r}
#CPUE units, SD and error structure
stocksynthesis.data$CPUEinfo

#Number of length bins
stocksynthesis.data$N_lbins

#Length bin method
stocksynthesis.data$Lbin_method

#Vector of age bins
stocksynthesis.data$agebin_vector

@@

```{r}
results <- get(load(atlantis_output_name))
devtools::load_all()
species=c(species)

@@

#Efficiency of the survey
effic <- data.frame(species=species, efficiency=eff_case)

# Assume uniform selectivity
sel<- data.frame(species=rep(species,length(age_classes)),
               agecl=age_classes,
              selex=sel_by_age)

#Sample survey - 3rd timestep simulates a summer survey
survey_out <- create_survey(dat=results$nums, time=survey_sample_full, species=species, boxes=boxes, effic=effic, selex=sel)

#Set effective sample size for age compositions
effN <- surveyEffN
highEffN <- data.frame(species=species, effN=rep(effN, length(species)))

@@

#Extract length composition data
ss_length_stdsurv <- calc_age2length(structn = ss_structnstd,
                                     resn = ss_resnstd,
                                     nums = age_comp_data,
                                     biolprm = results$biolprm, fgs = results$fgs,
                                     CVlenage = CVs$lenage, remove.zeroes=TRUE)

@@

# CV function
cv <- data.frame(species="Pacific_sardine", cv=CVs$survey)

@@

```{r}
catch_numbers <-  create_fishery_subset(dat = results$catch,
                                         time = fish_years,
                                         species = species,
                                         boxes = boxes)

#Sample catch in numbers
catch_total <- catch_numbers %>%
  group_by(time) %>%
  summarise(catch=sum(atoutput))

#Test writing CPUE in biomass
  stocksynthesis.data <- SS_write_ts(ss_data_list = stocksynthesis.data,
                ts_data = list(survObsBiom$atoutput[fish_years],
                  catch_total$catch[fish_years]),
                CVs = c(CVs$survey,
                        CVs$fishery),
                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             catch_total$time[fish_years]),
            sampling_month = list(rep(survey_month,nyears),
                                rep(fishing_month,nyears)),
                units = c("biomass","numbers"),
                data_type=c("CPUE","catch"),
                fleets = c(2,1))

@@

#Get the age bins
age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
age_bins <- sub("a","",age_bin_names)

@@

## Write age composition data for survey
stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data,
               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
               data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr+nyears-1)),
               sampling_month = list(rep(survey_month,nyears)),
               data_type = c("agecomp"),
               fleet_number = c(1),
               bins = list(age_bins),
               caal_bool = c(FALSE))
stocksynthesis.data$agecomp

#We end up using CAAL for the survey below, so let's generate fishery age comps instead
effN <- data.frame(species=species, effN=effN)

catch_numssshigh <- sample_fish(catch_numbers, effN)

@@

catch_length <- calc_age2length(structn = catch_structnss,
                                 resn = catch_resnss,
                                 nums = catch_numssshigh,
                                 biolprm = results$biolprm, fgs = results$fgs,
                                 maxbin = maxbin,
                                 CVlenage = CVs$lenage, remove.zeroes=TRUE)

@@

## Setting up EM
For the EM, we want conditional age-at-length (CAAL) and length composition data only. So we use the same function with our matrices of length and CAAL comp data to write to the data file.

@@

apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), rep(survey_month, nrow(comp_list[[2]])), rep(fishing_month,nrow(comp_list[[3]])),rep(fishing_month,nrow(comp_list[[4]])))

# Write CAAL and length composition data
stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data, comp_matrix = comp_list, data_rows = list((comp_list[[1]]$time-survey_sample_time)/timestep + 1 ,(survey_years-survey_sample_time)/timestep + 1,fish_years,fish_years),
        sampling_month = apply_month, data_type = rep(c("agecomp", "lencomp"),2), fleet_number = c(2,2,1,1),  bins = list(age_bins, length_bins, age_bins, length_bins), caal_bool = c(TRUE, rep(FALSE,3)))

head(stocksynthesis.data$lencomp)
head(stocksynthesis.data$agecomp)

#Change length bin structure to match atlantis data
stocksynthesis.data$lbin_vector <- length_bins

#Set lbin_method to 1 - this makes the population length bins match the data bins 
#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
stocksynthesis.data$lbin_method <- 1
stocksynthesis.data$binwidth <- "#"
stocksynthesis.data$minimum_size <- "#"
stocksynthesis.data$maximum_size <- "#"

#Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
stocksynthesis.data$age_info$minsamplesize <- rep(0.001,2)

SS_writedat_3.30(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
datfile_name),
                 overwrite=TRUE)
```

@@

len_nonburn <- ss_length_stdsurv$mulen %>%
  filter(time > burnin*timestep)
plot(len_nonburn$atoutput~len_nonburn$agecl, col=len_nonburn$time)

length.data <- data.frame("Year"=(len_nonburn$time-survey_sample_time)/timestep+1, length=len_nonburn$atoutput, Weight=NA, Sex="Female", age=as.integer(len_nonburn$agecl))

#require(bbmle)
#vb_estimates <-sample_fit_vbgf(length.data, 25, 45, 0.4,
#  0.1, 0.1, 20, 40, 0.05, 0.01, 0.01,
#  30, 50, 0.6, 0.3, 0.3, 0.5, 10)

@@

## Running stock synthesis
After the data and control file modifications, ensure you copy the local SS 3.3 executable into the folder containing the model files. Then run the following chunk.

```{r}
#Run stock synthesis model
run_stocksynthesis(species = species, model_dir = model_dir, show_output = TRUE)
